---
title: "final project"
author: "Ayush, Yuxiao Rui, Sirui Xu, Xia Wu"
date: '2022-11-29'
output:
  html_document:
    toc: yes
    toc_depth: 5
    toc_float: yes
    fig_height: 5
    fig_width: 7
  pdat_document:
    fig_height: 3.5
    fig_width: 3.5
  word_document:
    toc: no
  pdf_document:
    toc: no
    toc_depth: '5'
---

```{r,echo=FALSE}
knitr::opts_chunk$set(cache=TRUE, error=FALSE, message=FALSE, warning=FALSE,
                      echo=T,tidy=TRUE)
library(tidyverse)
library(here)
library(tidyr)
library(amap)
library(readxl)
library(ggplot2)
library(tableone)
library(ggalluvial)
```

```{r}
# load data
##dat <- read_xlsx("C:/Users/Lenovo/Desktop/Cornell Fall/Biostatistics/Project/sarcoma.xlsx", na = "NA")
#dat <- dat[complete.cases(dat),]

# load data
setwd("C:/Users/Lenovo/Desktop/Cornell Fall/Data Science/Biostats_Project")
dat <- read_xlsx("sarcoma.xlsx", na = "NA")
dat <- dat[complete.cases(dat),]

sarcoma <-  read_xlsx("sarcoma.xlsx", na = "NA") 
sarcoma <- sarcoma[complete.cases(sarcoma),]

```

# Problem 1:  What are the descriptive characteristics of our sample?  

```{r}
# Creating TableOne to understand the statistics of each parameter in the data

CreateTableOne(strata = "ARMCD", data = dat)
```

```{r}
plot_data <- dat %>% select (ID,ARMCD,STAGE,BESTRESP) %>%
             group_by(ARMCD,STAGE,BESTRESP) %>%
             summarise(COUNT = n()) %>% 
             filter(STAGE != 'NA') %>% filter(BESTRESP!= 'NA',
                                              BESTRESP!= 'NE') %>%
             mutate(BESTRESP = recode(BESTRESP, 'PD'= 'Progressive Disease', 'PR' = 'Partial Response', 
                                      'CR' = 'Complete Response', 'SD' = 'Stable Disease', 'NE' =
                                      'Inevaluable'),
                    ARMCD = recode(ARMCD, 'A' = 'DOX + EVO', 'B' = 'DOX'))


 ggplot(plot_data,
       aes(axis1 = STAGE,
           axis2 = BESTRESP,
           y = COUNT)) +
  geom_alluvium(aes(fill = ARMCD),curve_type = "cubic") +
  geom_stratum() +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)),size= 2.5) +
  theme_void() +
labs(title ="Participant Progression towards Response") +
theme(plot.title = element_text(hjust = 0.5)) +
guides(fill=guide_legend(title="Arms"))
```

```{r}
sarcoma %>% 
  filter(BESTRESP != "NA") %>% 
  ggplot(aes(BESTRESP, AGE, fill=BESTRESP)) + geom_boxplot() + theme_bw() + 
  stat_summary(fun='mean',geom='point', shape=23, size=3, fill='white') + 
  theme(legend.position = "none",plot.title = element_text(hjust = 0.5)) +
  scale_x_discrete(labels=c('Complete \nresponse', 'Progressive \ndisease', 'Partial \nresponse', 'Stable \ndisease')) + labs(x='', title='Best Response groups by age')
```



# Problem 2

## Problem 2.1：Are there any differences in baseline characteristics between the two arms of the trial?

### Problem 2.1: Check SEX
check if 2 arms have the same amount of male & female in different stage 
H0: The 2 arms have the same amount of male & female in different stage 
H1: The 2 arms do not have the same amount of male & female in different stage 
```{r}
prop.table(table(dat$SEX, dat$ARMCD))

ggplot(dat, aes(ARMCD, fill = SEX)) + 
  geom_bar(stat = "count")


# from ggplot, no obvious differences in sex between the two arms of the trial

results <- chisq.test(table(dat$SEX, dat$ARMCD))
results
# from chisq.test, p>0.05, fail to reject null. The 2 arms have the same amount of male & female in different stage 
```

### Problem 2.1: Check RACE
check if 2 arms have the same amount of RACE in different stage 
H0: The 2 arms have the same amount of RACE in different stage 
H1: The 2 arms do not have the same amount of RACE in different stage 
```{r}
prop.table(table(dat$RACE, dat$ARMCD))

ggplot(dat, aes(ARMCD, fill = RACE)) + 
  geom_bar(stat = "count")

# from the plot, Arm A does not have the race of AMERICAN INDIAN OR ALASKA NATIV, also otherraces distribution are different between the two arms of the trial

fisher.test(table(dat$RACE, dat$ARMCD), simulate.p.value = TRUE)
# from the fisher test, the p-value is greater than 0.05, so fail to reject null. The 2 arms have the same amount of RACE in different stage 
```
### Problem 2.1: Check ETHNIC

```{r}
prop.table(table(dat$ETHNIC, dat$ARMCD))

ggplot(dat, aes(ARMCD, fill = ETHNIC)) + 
  geom_bar(stat = "count")

# from the ggplot, no obvious differences in ETHNIC  between the two arms of the trial

results <- chisq.test(table(dat$ETHNIC, dat$ARMCD))


# from chisq.test, p>0.05, fail to reject null. The 2 arms have the same amount of ETHNIC in different stage 

```


### Problem 2.1: Check baseline BECOG
```{r}
prop.table(table(dat$BECOG, dat$ARMCD))

ggplot(dat, aes(ARMCD, fill = BECOG)) + 
  geom_bar(stat = "count")

# from the ggplot, no obvious differences in BECOG  between the two arms of the trial


fisher.test(table(dat$BECOG, dat$ARMCD), simulate.p.value = TRUE)
# from the fisher test, the p-value is greater than 0.05, so fail to reject null. The 2 arms have the same amount of BECOG in different stage 
```




## Problem 2.2：Are stage at diagnosis and histological grade at diagnosis different between the two groups? 
### Problem 2.2.a. STAGE $ ARMCD: 
check if 2 arms have the same amount of patients in different stage 
H0: The 2 arms have the same amount of patients in different stage 
H1: The 2 arms do not have the same amount of patients in different stage 
```{r}
prop.table(table(dat$STAGE, dat$ARMCD))

ggplot(dat, aes(ARMCD, fill = STAGE)) + 
  geom_bar(stat = "count")

results <- chisq.test(table(dat$STAGE, dat$ARMCD))
results
# chi-square result shows p<0.05，reject H0，so 2 arms do not have the same amount of patients in different stage 
```

### Problem 2.2.b. HISTGRDE $ ARMCD: 
check if 2 arms have the same amount of patients in each histological grade at diagnosis
H0: The 2 arms have the same amount of patients in each histological grade at diagnosis
H1: The 2 arms do not have the same amount of patients in each histological grade at diagnosis
```{r}
prop.table(table(dat$HISTGRDE, dat$ARMCD))

ggplot(dat, aes(ARMCD, fill = HISTGRDE)) + 
  geom_bar(stat = "count")

results <- chisq.test(table(dat$HISTGRDE, dat$ARMCD))
results$expected
# The expected count of Low Grade is smaller than 5，so chi-square is not suitable. We choose fisher test.
# fisher
fisher.test(table(dat$HISTGRDE, dat$ARMCD), simulate.p.value = TRUE)

# p>0.05，fail to reject H0，so 2 arms do not have the same amount of patients in each histological grade at diagnosis
```


## Problem 2.3： Is there a difference in outcomes for patients in different arms of the trial?
### Problem 2.3.a: Check if the tumor size change has a difference for patients in different arms of the trial
H0: There is no difference in tumor size change for patients in different arms of the trial
H1: There is a difference in tumor size change for patients in different arms of the trial
```{r}
# first define the tumor_size_change
dat <- dat %>%
  mutate(
    tumor_size_change = 100*(as.numeric(EOT_SUM_DIAMETER)-as.numeric(BL_SUM_DIAMETER))/(as.numeric(BL_SUM_DIAMETER)))

# check if the distribution is normal, so we can decide to use t-test or wilcox test
ggplot(dat, aes(tumor_size_change, fill = ARMCD)) + 
  geom_density(alpha=0.5)
shapiro.test(dat$tumor_size_change[dat$ARMCD == "A"])
shapiro.test(dat$tumor_size_change[dat$ARMCD == "B"])
# p<0.05，we reject null and states the data is not normal, we cannot use t test. Then we use wilcox.

wilcox.test(tumor_size_change~ARMCD, dat)
# p>0.05，fail to reject H0. There is no difference in tumor size change for patients in different arms of the trial
```


### Problem 2.3.b: Check if the best response has a difference for patients in different arms of the trial
H0: There is no association in best response for patients in different arms of the trial
H1: There is an association in best response for patients in different arms of the trial
```{r}
# set the response variable BESTRESP to be 0 or 1, (no tumor response or with tumor response)
dat <- dat %>%
  mutate(
    BESTRESP = recode(as.character(BESTRESP), "PD" = "0", "SD" = "0", "NE" = "0", 
                      "PR" = "1", "CR" = "1")
  )

prop.table(table(dat$BESTRESP, dat$ARMCD))

ggplot(dat, aes(ARMCD, fill = BESTRESP)) + 
  geom_bar(stat = "count")

results <- chisq.test(table(dat$BESTRESP, dat$ARMCD))
results
# chi-square result shows p<0.05，reject H0. There is no association in best response for patients in different arms of the trial

```


# Problem 3
## Problem 3.1：What factors are associated with tumor response (binary)?

```{r}
# clean the data
dat$EOT_SUM_DIAMETER=as.numeric(dat$EOT_SUM_DIAMETER)

# drop NA
dat=dat %>% filter(BL_SUM_DIAMETER!="NA") %>% filter(EOT_SUM_DIAMETER!="NA")
dat=dat %>% filter(BESTRESP!="NA")  %>%
  mutate(
    age_group= ifelse(AGE > 60, "older than 60", "younger or equal to 60") )

dat_young=dat%>%filter(age_group=="younger or equal to 60")
dat_old=dat%>%filter(age_group=="older than 60")



# change y to numeric
dat$BESTRESP = as.numeric(dat$BESTRESP)


# in arm A, treatments include DOX+EVO
# in arm B, treatments include DOX only
mylogit <- glm(dat$BESTRESP ~ dat$ARMCD + age_group +RACE + ETHNIC, data = dat, family = "binomial")
summary(mylogit)

mylogit$coefficients %>% exp()

# interpretation
# For ARMCDB, AGE, and RACEBLACK OR AFRICAN AMERICAN, the p-value is significant, so they are  associated with tumor response; and we can interprete them as 

# If a patient belongs to arm B, it will decrease the log odds of getting a tumor response by 0.549;so arm A (DOX+EVO) gives more tumor response
# Each 1 unit increase in AGE will decrease the log odds of getting a tumor response by 0.016
# If a patient's race is BLACK OR AFRICAN AMERICAN, it will decrease the log odds of getting a tumor response by 3.20879
```

## Alternate Model

```{r}
# clean the data
dat$EOT_SUM_DIAMETER=as.numeric(dat$EOT_SUM_DIAMETER)

# drop NA
dat=dat %>% filter(BL_SUM_DIAMETER!="NA") %>% filter(EOT_SUM_DIAMETER!="NA")
dat=dat %>% filter(BESTRESP!="NA")  %>%
  mutate(
    age_group= ifelse(AGE > 60, "older than 60", "younger or equal to 60") )

dat_young=dat%>%filter(age_group=="younger or equal to 60")
dat_old=dat%>%filter(age_group=="older than 60")



# change y to numeric
dat$BESTRESP = as.numeric(dat$BESTRESP)
dat$HISTGRP = as.factor(dat$HISTGRP)
dat$STAGE = as.factor(dat$STAGE)
dat$PRADFL = as.factor(dat$PRADFL)
dat$PCHEMFL = as.factor(dat$PCHEMFL)



# in arm A, treatments include DOX+EVO
# in arm B, treatments include DOX only
mylogit <- glm(dat$BESTRESP ~ dat$ARMCD + age_group +RACE + ETHNIC + HISTGRP + STAGE + PRADFL + PCHEMFL, data = dat, family = "binomial")
summary(mylogit)

mylogit$coefficients %>% exp()

# interpretation
# For ARMCDB, AGE, and RACEBLACK OR AFRICAN AMERICAN, the p-value is significant, so they are  associated with tumor response; and we can interprete them as 

# If a patient belongs to arm B, it will decrease the log odds of getting a tumor response by 0.549;so arm A (DOX+EVO) gives more tumor response
# Each 1 unit increase in AGE will decrease the log odds of getting a tumor response by 0.016
# If a patient's race is BLACK OR AFRICAN AMERICAN, it will decrease the log odds of getting a tumor response by 3.20879
```


## Problem 3.2：What factors are associated With percent change in sum of tumor diameters (continuous)? 
```{r}
model2 <- lm(dat$tumor_size_change ~ dat$ARMCD + AGE+RACE + ETHNIC, data = dat)
summary(model2)

# no significant factor
```


# Problem 4
## Problem 4.1: What conclusions can you make about the treatments’ effects on tumor response and percent change in sum of tumor diameters? 


treatments’ effects on tumor response: According to the logistic regression function we did in Problem 3.1, if a patient belongs to arm B, it will decrease the log odds of getting a tumor response by 0.549; so arm A (DOX+EVO) gives more tumor response.

treatments’ effects on percent change in sum of tumor diameters: According to the linear regression function we did in Problem 3.2, there is no significant factor are associated with percent change in sum of tumor diameters.




## Problem 4.2:Potential differential treatment effects in subgroups of patients (e.g., gender, race, age groups, etc.)

### for subgroup SEX, if treatment allocation (ARMCD) cause a change on the outcome (Best response & tumor size change)
```{r}
# split data by sex
dat_female=dat%>%filter(SEX=="F")
dat_male=dat%>%filter(SEX=="M")

# check treatment effects in Best response
female_logit <- glm(dat_female$BESTRESP ~ dat_female$ARMCD, data = dat_female, family = "binomial")
summary(female_logit)
# for female, the p-value of the treatment allocation (ARMCD) is not significant, so ARMCD does not cause a change on the tumor response in female. 

male_logit <- glm(dat_male$BESTRESP ~ dat_male$ARMCD, data = dat_male, family = "binomial")
summary(male_logit)
# for male, the p-value of the treatment allocation (ARMCD) is significant, so ARMCD does cause a change on the tumor response in male. If a male patient belongs to arm B, it will decrease the log odds of getting a tumor response by 0.61.


# check treatment effects in tumor size change 
female_lm <- lm(dat_female$tumor_size_change ~ dat_female$ARMCD, data = dat_female)
summary(female_lm)
# for female, the p-value of the treatment allocation (ARMCD) is not significant, so ARMCD does not cause a change on the tumor size in female. 


male_lm <- lm(dat_male$tumor_size_change ~ dat_male$ARMCD, data = dat_male)
summary(male_lm)
# For male,the p-value of the treatment allocation (ARMCD) is not significant, so ARMCD does not cause a change on the tumor size in female. 
```

### for subgroup RACE, if treatment allocation (ARMCD) cause a change on the outcome (Best response & tumor size change)
```{r}
# split data by the top 3 representative race
dat_white=dat%>%filter(RACE=="WHITE")
dat_black=dat%>%filter(RACE=="BLACK OR AFRICAN AMERICAN")
dat_asia=dat%>%filter(RACE=="ASIAN")


# check treatment effects in Best response
white_logit <- glm(BESTRESP ~ ARMCD, data = dat_white, family = "binomial")
summary(white_logit)
# for white race, the p-value of the treatment allocation (ARMCD) is significant, so ARMCD does cause a change on the tumor response in female. 
# If a white patient belongs to arm B, it will decrease the log odds of getting a tumor response by 0.51

black_logit <- glm(BESTRESP ~ ARMCD, data = dat_black, family = "binomial")
summary(black_logit)
# for black race, the p-value of the treatment allocation (ARMCD) is not significant, so ARMCD does not cause a change on the tumor response in black race.

asia_logit <- glm(BESTRESP ~ ARMCD, data = dat_asia, family = "binomial")
summary(asia_logit)
# for Asian race, the p-value of the treatment allocation (ARMCD) is not significant, so ARMCD does not cause a change on the tumor response in asia race.

# check treatment effects in tumor size change 
white_logit <- lm(tumor_size_change ~ ARMCD, data = dat_white)
summary(white_logit)
# for white race, the p-value of the treatment allocation (ARMCD) is not significant, so ARMCD does not cause a change on the tumor size in white race.

black_logit <- lm(tumor_size_change ~ ARMCD, data = dat_black)
summary(black_logit)
# for black race, the p-value of the treatment allocation (ARMCD) is not significant, so ARMCD does not cause a change on the tumor size in balck race.

asia_logit <- lm(tumor_size_change ~ ARMCD, data = dat_asia)
summary(asia_logit)
# for black race, the p-value of the treatment allocation (ARMCD) is not significant, so ARMCD does not cause a change on the tumor size in asian race.

```

### for subgroup Age groups, if treatment allocation (ARMCD) cause a change on the outcome (Best response & tumor size change)
```{r}
# split the data by 2 age groups
dat <- dat %>%
  mutate(
    age_group= ifelse(AGE > 60, "older than 60", "younger or equal to 60") )

dat_young=dat%>%filter(age_group=="younger or equal to 60")
dat_old=dat%>%filter(age_group=="older than 60")

# check treatment effects in Best response
young_logit <- glm(BESTRESP ~ ARMCD, data = dat_young, family = "binomial")
summary(young_logit)
# for people younger or equal to 60, the p-value of the treatment allocation (ARMCD) is significant, so ARMCD does cause a change on the tumor response. If a patient who is younger or equal to 60 and belongs to arm B, it will decrease the log odds of getting a tumor response by 0.78.

old_logit <- glm(BESTRESP ~ ARMCD, data = dat_old, family = "binomial")
summary(old_logit)
# for people older than 60 male, the p-value of the treatment allocation (ARMCD) is not significant, so ARMCD does not cause a change on the tumor response in older people.


# check treatment effects in tumor size change 
young_lm <- lm(tumor_size_change ~ ARMCD, data = dat_young)
summary(young_lm)
# for people younger or equal to 60, the p-value of the treatment allocation (ARMCD) is not significant, so ARMCD does not cause a change on the tumor size in female. 


old_lm <- lm(tumor_size_change ~ ARMCD, data = dat_old)
summary(old_lm)
# For people older than 60,the p-value of the treatment allocation (ARMCD) is not significant, so ARMCD does not cause a change on the tumor size in female. 
```


