---
title: "final project"
author: "Ayush, Yuxiao Rui, Sirui Xu, Xia Wu"
date: '2022-11-29'
output:
  html_document:
    toc: yes
    toc_depth: 5
    toc_float: yes
    fig_height: 5
    fig_width: 7
  pdat_document:
    fig_height: 3.5
    fig_width: 3.5
  word_document:
    toc: no
---

```{r,echo=FALSE}
knitr::opts_chunk$set(cache=TRUE, error=FALSE, message=FALSE, warning=FALSE,
                      echo=T, tidy.opts=list(width.cutoff=60),tidy=TRUE)
library(tidyverse)
library(here)
library(tidyr)
library(amap)
library(readxl)
library(ggplot2)

# load data
dat <- read_xlsx("C:/Users/Lenovo/Desktop/Cornell Fall/Biostatistics/Project/sarcoma.xlsx", na = "NA")
dat <- dat[complete.cases(dat),]
```

# Problem 2

## Problem 2.1：Are there any differences in baseline characteristics between the two arms of the trial?

### Problem 2.1: Check SEX
check if 2 arms have the same amount of male & female in different stage 
H0: The 2 arms have the same amount of male & female in different stage 
H1: The 2 arms do not have the same amount of male & female in different stage 
```{r}
prop.table(table(dat$SEX, dat$ARMCD))

ggplot(dat, aes(ARMCD, fill = SEX)) + 
  geom_bar(stat = "count")





# from ggplot, no obvious differences in sex between the two arms of the trial

results <- chisq.test(table(dat$SEX, dat$ARMCD))
# from chisq.test, p>0.05, fail to reject null. The 2 arms have the same amount of male & female in different stage 
```
### Problem 2.1: Check RACE
check if 2 arms have the same amount of RACE in different stage 
H0: The 2 arms have the same amount of RACE in different stage 
H1: The 2 arms do not have the same amount of RACE in different stage 
```{r}
prop.table(table(dat$RACE, dat$ARMCD))

ggplot(dat, aes(ARMCD, fill = RACE)) + 
  geom_bar(stat = "count")

# from the plot, Arm A does not have the race of AMERICAN INDIAN OR ALASKA NATIV, also otherraces distribution are different between the two arms of the trial

fisher.test(table(dat$RACE, dat$ARMCD), simulate.p.value = TRUE)
# from the fisher test, the p-value is greater than 0.05, so fail to reject null. The 2 arms have the same amount of RACE in different stage 
```
### Problem 2.1: Check ETHNIC

```{r}
prop.table(table(dat$ETHNIC, dat$ARMCD))

ggplot(dat, aes(ARMCD, fill = ETHNIC)) + 
  geom_bar(stat = "count")

# from the ggplot, no obvious differences in ETHNIC  between the two arms of the trial

results <- chisq.test(table(dat$ETHNIC, dat$ARMCD))


# from chisq.test, p>0.05, fail to reject null. The 2 arms have the same amount of ETHNIC in different stage 

```


### Problem 2.1: Check baseline BECOG
```{r}
prop.table(table(dat$BECOG, dat$ARMCD))

ggplot(dat, aes(ARMCD, fill = BECOG)) + 
  geom_bar(stat = "count")

# from the ggplot, no obvious differences in BECOG  between the two arms of the trial


fisher.test(table(dat$BECOG, dat$ARMCD), simulate.p.value = TRUE)
# from the fisher test, the p-value is greater than 0.05, so fail to reject null. The 2 arms have the same amount of BECOG in different stage 
```




## Problem 2.2：Are stage at diagnosis and histological grade at diagnosis different between the two groups? 
### Problem 2.2.a. STAGE $ ARMCD: 
check if 2 arms have the same amount of patients in different stage 
H0: The 2 arms have the same amount of patients in different stage 
H1: The 2 arms do not have the same amount of patients in different stage 
```{r}
prop.table(table(dat$STAGE, dat$ARMCD))

ggplot(dat, aes(ARMCD, fill = STAGE)) + 
  geom_bar(stat = "count")

results <- chisq.test(table(dat$STAGE, dat$ARMCD))
results
# chi-square result shows p<0.05，reject H0，so 2 arms do not have the same amount of patients in different stage 
```

### Problem 2.2.b. HISTGRDE $ ARMCD: 
check if 2 arms have the same amount of patients in each histological grade at diagnosis
H0: The 2 arms have the same amount of patients in each histological grade at diagnosis
H1: The 2 arms do not have the same amount of patients in each histological grade at diagnosis
```{r}
prop.table(table(dat$HISTGRDE, dat$ARMCD))

ggplot(dat, aes(ARMCD, fill = HISTGRDE)) + 
  geom_bar(stat = "count")

results <- chisq.test(table(dat$HISTGRDE, dat$ARMCD))
results$expected
# The expected count of Low Grade is smaller than 5，so chi-square is not suitable. We choose fisher test.
# fisher
fisher.test(table(dat$HISTGRDE, dat$ARMCD), simulate.p.value = TRUE)

# p>0.05，fail to reject H0，so 2 arms do not have the same amount of patients in each histological grade at diagnosis
```


## Problem 2.3： Is there a difference in outcomes for patients in different arms of the trial?
### Problem 2.3.a: Check if the tumer size change has a difference for patients in different arms of the trial
H0: There is no difference in tumer size change for patients in different arms of the trial
H1: There is a difference in tumer size change for patients in different arms of the trial
```{r}
# first define the tumer_size_change
dat <- dat %>%
  mutate(
    tumer_size_change = 100*(as.numeric(EOT_SUM_DIAMETER)-as.numeric(BL_SUM_DIAMETER))/(as.numeric(BL_SUM_DIAMETER)))

# check if the distribution is normal, so we can decide to use t-test or wilcox test
ggplot(dat, aes(tumer_size_change, fill = ARMCD)) + 
  geom_density()
shapiro.test(dat$tumer_size_change[dat$ARMCD == "A"])
shapiro.test(dat$tumer_size_change[dat$ARMCD == "B"])
# p<0.05，we reject null and states the data is not normal, we cannot use t test. Then we use wilcox.

wilcox.test(tumer_size_change~ARMCD, dat)
# p>0.05，fail to reject H0. There is no difference in tumer size change for patients in different arms of the trial
```


### Problem 2.3.b: Check if the best response has a difference for patients in different arms of the trial
H0: There is no association in best response for patients in different arms of the trial
H1: There is an association in best response for patients in different arms of the trial
```{r}
# set the response variable BESTRESP to be 0 or 1, (no tumer response or with tumer response)
dat <- dat %>%
  mutate(
    BESTRESP = recode(as.character(BESTRESP), "PD" = "0", "SD" = "0", "NE" = "0", 
                      "PR" = "1", "CR" = "1")
  )

prop.table(table(dat$BESTRESP, dat$ARMCD))

ggplot(dat, aes(ARMCD, fill = BESTRESP)) + 
  geom_bar(stat = "count")

results <- chisq.test(table(dat$BESTRESP, dat$ARMCD))
# chi-square result shows p<0.05，reject H0. There is no association in best response for patients in different arms of the trial

```



## Problem 3.1：What factors are associated with tumor response (binary)?

```{r}
# clean the data
dat$EOT_SUM_DIAMETER=as.numeric(dat$EOT_SUM_DIAMETER)

# drop NA
dat=dat %>% filter(BL_SUM_DIAMETER!="NA") %>% filter(EOT_SUM_DIAMETER!="NA")
dat=dat %>% filter(BESTRESP!="NA")



# change y to numeric
dat$BESTRESP = as.numeric(dat$BESTRESP)


# in arm A, treatments include DOX+EVO
# in arm B, treatments include DOX only
mylogit <- glm(dat$BESTRESP ~ dat$ARMCD + AGE+RACE + ETHNIC, data = dat, family = "binomial")
summary(mylogit)

# interpretation
# For ARMCDB, AGE, and RACEBLACK OR AFRICAN AMERICAN, the p-value is significant, so they are  associated with tumor response; and we can interprete them as 

# If a patient belongs to arm B, it will decrease the log odds of getting a tumer response by 0.549;so arm A (DOX+EVO) gives more tumer response
# Each 1 unit increase in AGE will decrease the log odds of getting a tumer response by 0.016
# If a patient's race is BLACK OR AFRICAN AMERICAN, it will decrease the log odds of getting a tumer response by 3.20879
```


## Problem 3.2：What factors are associated With percent change in sum of tumor diameters (continuous)? 
```{r}
model2 <- lm(dat$tumer_size_change ~ dat$ARMCD + AGE+RACE + ETHNIC, data = dat)
summary(model2)
# no significant factor
```

