---
title: "Final Project"
author: "Ayush Bhattacharya, Yuxiao Rui, Sirui Xu, Xia Wu"
date: '2022-11-29'
output:
  html_document:
    toc: yes
    toc_depth: 5
    toc_float: yes
    fig_height: 5
    fig_width: 7
  pdat_document:
    fig_height: 3.5
    fig_width: 3.5
  word_document:
    toc: no
  pdf_document:
    toc: no
    toc_depth: '5'
---

```{r,echo=FALSE,warning=FALSE, include=FALSE}
knitr::opts_chunk$set(cache=TRUE, error=FALSE, message=FALSE, warning=FALSE,
                      echo=T,tidy=TRUE)
library(tidyverse)
library(here)
library(tidyr)
library(amap)
library(readxl)
library(ggplot2)
library(tableone)
library(ggalluvial)
library(ggpubr)
```

### 1. Introduction

Soft tissue sarcomas (STS) is a general term for tumors that begin in soft tissues. These tumors can form anywhere in the body but are most common in the legs, chest, and arms (NIH, 2016) . A typical treatment for STS is the doxorubicin (DOX) treatment (Verschoor, 2020).

Nowadays, researchers are trying to improve patient outcomes by testing different treatments. A randomized trial was done to determine if evofosfamide (EVO) could improve outcomes in individuals with STS. This trial has two arms: in arm A, patients were treated with both DOX and EVO, and in arm B, patients were treated with DOX alone.

The primary research question is whether the tumor response is improved for patients receiving DOX + EVO treatment compared to DOX treatment itself. Additionally, we aim to investigate if the characterististics of patients in the two arms are similar. The general approaches employed in the statistical analysis are regression models including both linear regression and logistic regression.

Lastly, the data was provided by sarcoma experts and members of the Sarcoma Alliance for Research through Collaboration (SARC). Histological grade at diagnosis.


### 2. Methodology

The dataset was obtained between Sept 26, 2011, and Jan 22, 2014. In a randomized control trial, 640 patients were randomly assigned to a treatment group (317 received DOX + EVO, while 323 received DOX alone). They included Patients 15 years or older with a diagnosis of an advanced unresectable or metastatic soft-tissue sarcoma, of intermediate or high grade, for which no standard curative therapy was available. Patients who had received previous systemic therapy for advanced or other certain diseases were excluded.

Before we conducted the statistical analysis, we performed data cleansing, normality checks, and other exploratory data analysis utilizing TableOne, data visualizations (e.g. bar chart), and the Chi-square test and Fisher Exact Test for checking the differences among patients in terms of characteristics and conditions. 

Then we decided to use both logistic and linear regression to evaluate the tumor response for each group. 
To investigate what factors are associated with tumor response, we implemented a logistic regression model as the response variable tumor response (BESRESP) is a binary variable. It has either “0” (no tumor response ) or “1” (with tumor response). To investigate what factors are associated with percent change in the sum of tumor diameters, we used a linear regression model as the response variable tumor_size_change is continuous. 


### 3. Analysis 

```{r, echo=FALSE}
# Loading Data

dat <- read_xlsx("sarcoma.xlsx", na = "NA")
dat <- dat[complete.cases(dat),]

sarcoma <-  read_xlsx("sarcoma.xlsx", na = "NA") 
sarcoma <- sarcoma[complete.cases(sarcoma),]
```

#### 3.1. What are the descriptive characteristics of our sample?  

```{r, echo=FALSE}
# Cleaning Data
dat_grade <- dat %>%
  mutate(
    HISTGRDE = recode(as.character(HISTGRDE), 
                      "Grade 4" = "High Grade", 
                      "Grade 2" = "Intermediate Grade", 
                      "Grade 3" = "Intermediate Grade", 
                      "Grade 2/3" = "Intermediate Grade", 
                      "Grade 1" = "Low Grade")
  )

# Creating TableOne to understand the statistics of each parameter in the data
CreateTableOne(strata = "ARMCD", data = dat_grade)
```

TableOne analysis was carried out to understand the data variables and look for statistical significance within groups of data. We see that stage of diagnosis comes out as a factor of statistical significance with a p-value of <0.05.

```{r, echo=FALSE, warning=FALSE}
## Alluvial Plot for Patient Progression

plot_data <- dat %>% select (ID,ARMCD,STAGE,BESTRESP) %>%
             group_by(ARMCD,STAGE,BESTRESP) %>%
             summarise(COUNT = n()) %>% 
             filter(STAGE != 'NA') %>% filter(BESTRESP!= 'NA',
                                              BESTRESP!= 'NE') %>%
             mutate(BESTRESP = recode(BESTRESP, 'PD'= 'Progressive Disease', 'PR' = 'Partial Response', 
                                      'CR' = 'Complete Response', 'SD' = 'Stable Disease', 'NE' =
                                      'Inevaluable'),
                    ARMCD = recode(ARMCD, 'A' = 'DOX + EVO', 'B' = 'DOX'))


                        ggplot(plot_data,
                         aes(axis1 = STAGE,
                             axis2 = BESTRESP,
                             y = COUNT)) +
                        geom_alluvium(aes(fill = ARMCD),curve_type = "cubic") +
                        geom_stratum() +
                        geom_text(stat = "stratum", aes(label = after_stat(stratum)),size= 2.5) +
                        theme_void() +
                        labs(title ="Participant Progression towards Response") +
                        theme(plot.title = element_text(hjust = 0.5)) +
                        guides(fill=guide_legend(title="Arms"))
```

The alluvial plot shows the patient outcome patterns taking into consideration their initial diagnosis category. The comparison of treatment arms gives us a better visual understanding of the movement of patients who are assigned Arm A vs Arm B. One takeaway is that we see that more proportion of patients in Arm B move from Stage 3 to Stable disease - indicating a better positive outcome for patients given DOX + EVO.

```{r, echo=FALSE}
## Response Groups by Age

sarcoma %>% 
filter(BESTRESP != "NA") %>% 
ggplot(aes(BESTRESP, AGE, fill=BESTRESP)) + 
geom_boxplot() + 
theme_bw() + 
stat_summary(fun='mean',geom='point', shape=23, size=3, fill='white') + 
theme(legend.position = "none",
      plot.title = element_text(hjust = 0.5)) +
scale_x_discrete(labels=c('Complete \nresponse', 
                          'Progressive \ndisease', 
                          'Partial \nresponse', 
                          'Stable \ndisease')) + 
labs(x='', title='Best Response groups by age')
```

We see from the box plots that patients outcomes with respect to age give very similar results. 


#### 3.2. Are there any differences in baseline characteristics between the two arms of the trial?

##### 3.2.1 Checking Gender

check if the two arms have the same proportion of male & female in different arms 

H0: The two arms have the same proportion of male & female in different arms 
H1: The two arms do not have the same proportion of male & female in different arms 

```{r, echo=FALSE}
prop.table(table(dat$SEX, dat$ARMCD))

ggplot(dat, aes(ARMCD, fill = SEX)) + 
geom_bar(stat = "count")+  
theme_bw() +
labs(title = "Bar Plot of Gender in Two Arms", x="Treatment Arm", y= "Count") +
theme(plot.title = element_text(hjust = 0.5)) 

# from ggplot, no obvious differences in sex between the two arms of the trial

chisq.test(table(dat$SEX, dat$ARMCD))

# from chisq.test, p>0.05, fail to reject null. The 2 arms have the same amount of male & female in different stage 
```

The bar plot of the proportion of patients by gender shows that the number of males and females respectively in each arm is similar. Specifically, there are 151 females and 125 males in Arm A, and 149 females and 126 males in Arm B. 

From Pearson’s chi-squared test, the p-value is 0.9689, which is greater than the 0.05 significance level, so we fail to reject the null hypothesis. Therefore, we conclude that the two arms have the same proportion of males and females respectively in different arms. 

##### 3.2.2 Checking Race

check if two arms have the same proportion of race in different arms 

H0: The two arms have the same proportion of race in different arms 
H1: The two arms do not have the same proportion of race in different arms 

```{r, echo=FALSE}
prop.table(table(dat$RACE, dat$ARMCD))

ggplot(dat, aes(ARMCD, fill = RACE)) + 
geom_bar(stat = "count")+
theme_bw() +
labs(title = "Bar Plot of Race in Two Arms" ,x="Treatment Arm", y= "Count") +
theme(plot.title = element_text(hjust = 0.5))

# from the plot, Arm A does not have the race of AMERICAN INDIAN OR ALASKA NATIV, also other races distribution are different between the two arms of the trial

fisher.test(table(dat$RACE, dat$ARMCD), simulate.p.value = TRUE)
# from the fisher test, the p-value is greater than 0.05, so fail to reject null. The 2 arms have the same amount of RACE in different stage 
```

We generate a bar plot of the proportion of race distribution. White people are the majority of the patients in this study, which accounts for 82.16%. By fisher’s exact test, the p-value is 0.1919, greater than the significance level. Hence we accept the null hypothesis that both arms have the same proportion of races.

##### 3.2.3. Checking Ethnicity

check if two arms have the same proportion of ethnicity in different arms 

H0: The two arms have the same proportion of ethnicity in different arms 
H1: The two arms do not have the same proportion of ethnicity in different arms 

```{r, echo=FALSE}
prop.table(table(dat$ETHNIC, dat$ARMCD))

ggplot(dat, aes(ARMCD, fill = ETHNIC)) + 
geom_bar(stat = "count")+
theme_bw() +
labs(title = "Bar Plot of Ethnicity in Two Arms",  x="Treatment Arm", y= "Count") +
theme(plot.title = element_text(hjust = 0.5))

# from the ggplot, no obvious differences in ETHNIC  between the two arms of the trial

chisq.test(table(dat$ETHNIC, dat$ARMCD))


# from chisq.test, p>0.05, fail to reject null. The 2 arms have the same amount of ETHNIC in different stage 

```

We generate a bar plot of the proportion of ethnicity distribution. Majority of the patients are not hispanic or latino in this study. By Pearson's Chi-squared test, the p-value is 0.992, greater than the significance level. Hence we accept the null hypothesis that both arms have the same proportions of ethnicities.

#### 3.3. Are stage at diagnosis and histological grade at diagnosis different between the two groups? 

##### 3.3.1. STAGE & ARMCD

check if two arms have the same proportion of patients in different stages

H0: The two treatment arms have the same proportion of patients in different stages.
H1: The two treatment arms do not have the same proportion of patients in different stages.

```{r, echo=FALSE}
prop.table(table(dat$STAGE, dat$ARMCD))

ggplot(dat, aes(ARMCD, fill = STAGE)) + 
geom_bar(stat = "count") +
theme_bw() +
labs(title = "Bar Plot of Stage at Diagnosis in Two Arms",x="Treatment Arm", y= "Count") +
theme(plot.title = element_text(hjust = 0.5))

chisq.test(table(dat$STAGE, dat$ARMCD))

# chi-square result shows p<0.05，reject H0，so 2 arms do not have the same amount of patients in different stage 
```

To test if there is a difference in outcomes for patients in different arms of the trial, stage at diagnosis and best response are considered at first. In the bar plot of patients in each stage by arms, patients in arm A resulting in stage III is slightly more than those in arm B. However, patients in arm A resulting in other stages are slightly less than those in arm B. By Pearson's Chi-squared test, the p-value is 0.02, less than the significance level. Hence we accept the alternate hypothesis that the two treatment arms do not have the same proportion of patients in different stages.

##### 3.3.2. HISTGRDE & ARMCD
check if two arms have the same amount of patients in each histological grade at diagnosis

H0: The two arms have the same amount of patients in each histological grade at diagnosis
H1: The two arms do not have the same amount of patients in each histological grade at diagnosis

```{r, echo=FALSE}
dat_grade <- dat %>%
  mutate(
    HISTGRDE = recode(as.character(HISTGRDE), 
                      "Grade 4" = "High Grade", 
                      "Grade 2" = "Intermediate Grade", 
                      "Grade 3" = "Intermediate Grade", 
                      "Grade 2/3" = "Intermediate Grade", 
                      "Grade 1" = "Low Grade")
  )

prop.table(table(dat_grade$HISTGRDE, dat_grade$ARMCD))

dat_grade %>% ggplot(aes(y = HISTGRDE, fill = ARMCD)) +
geom_bar(stat = "count") +
theme_bw() +
labs(title = "Bar Plot of Histological Grade in 2 Arms",y="Histological Grade", x= "Count") +
theme(plot.title = element_text(hjust = 0.5))


results <- chisq.test(table(dat_grade$HISTGRDE, dat_grade$ARMCD))
results$expected
# The expected count of Low Grade is smaller than 5，so chi-square is not suitable. We choose fisher test.
# fisher
fisher.test(table(dat$HISTGRDE, dat$ARMCD), simulate.p.value = TRUE)

# p>0.05，fail to reject H0，so 2 arms do not have the same amount of patients in each histological grade at diagnosis
```

Under histological grade at diagnosis, all patients in 2 arms are divided into 3 groups, low grade, intermediate grade, and high grade. 

Grade 1 is defined as low, grade 4 is defined as high, and other 3 grades, including grade 2, grade 2 or 3, and grade 3, are defined as intermediate grades.

Most patients have intermediate histological grades at diagnosis, then have high grades. 
By using the fisher’s test, as the p-value is 0.4893 for this test, which is greater than the 0.05 significance level, we fail to reject the null hypothesis, leading to the conclusion that two arms have the same number of patients in each histological grade at diagnosis.

#### 3.4. Is there a difference in outcomes for patients in different arms of the trial?

##### 3.4.1. Tumor size change
Check if the tumor size change has a difference for patients in different arms of the trial

H0: There is no difference in tumor size change for patients in different arms of the trial
H1: There is a difference in tumor size change for patients in different arms of the trial

```{r, echo=FALSE}
# first define the tumor_size_change
dat <- dat %>%
  mutate(
    tumor_size_change = 100*(as.numeric(EOT_SUM_DIAMETER)-as.numeric(BL_SUM_DIAMETER))/(as.numeric(BL_SUM_DIAMETER)))

# check if the distribution is normal, so we can decide to use t-test or wilcox test
ggplot(dat, aes(tumor_size_change, fill = ARMCD)) + 
geom_density(alpha=0.5) +
theme_bw() +
labs(title = "Density Plot of Tumor Size Change in two Arms") +
theme(plot.title = element_text(hjust = 0.5))

shapiro.test(dat$tumor_size_change[dat$ARMCD == "A"])
shapiro.test(dat$tumor_size_change[dat$ARMCD == "B"])
# p<0.05，we reject null and states the data is not normal, we cannot use t test. Then we use wilcox.

wilcox.test(tumor_size_change~ARMCD, dat)
# p>0.05，fail to reject H0. There is no difference in tumor size change for patients in different arms of the trial
```

We check for normality in the data using Shapiro-Wilk normality test. We see that for outcomes of both arms the distribution is not normal. Hence we perform non-parametric Wilcoxon rank sum test to compare the tumor outcomes of the two different arms. 

As the p-value is greater than the 0.05 significance level, we accept the null hypothesis, leading to the conclusion that there is no difference in tumor size change for patients in different arms of the trial.

##### 3.4.2. Best response in 2 arms
Check if the best response has a difference for patients in different arms of the trial

H0: There is no association in best response for patients in different arms of the trial
H1: There is an association in best response for patients in different arms of the trial

```{r, echo=FALSE}
# set the response variable BESTRESP to be 0 or 1, (no tumor response or with tumor response)
dat <- dat %>%
  mutate(
    BESTRESP = recode(as.character(BESTRESP), "PD" = "0", "SD" = "0", "NE" = "0", 
                      "PR" = "1", "CR" = "1")
  )

prop.table(table(dat$BESTRESP, dat$ARMCD))

ggplot(dat, aes(ARMCD, fill = BESTRESP)) + 
geom_bar(stat = "count") +
theme_bw() +
labs(title = "Bar Plot of Best Response in Two Arms",x="Treatment Arm", y= "Count") +
theme(plot.title = element_text(hjust = 0.5))

chisq.test(table(dat$BESTRESP, dat$ARMCD))

# chi-square result shows p<0.05，reject H0. There is no association in best response for patients in different arms of the trial

```

To test if there is a difference in outcomes for patients in different arms of the trial, best response is considered. Patients are divided into two groups of best response, no tumor response (stable disease, progressive disease, and inevaluable) and tumor response (partial response and complete response). The number of people with tumor response in arm A is more than those in arm B. The p-values from Pearson’s chi-squared test for the test is smaller than the 0.05 significance level. Hence, we reject the null hypothesis and accept the alternate hypothesis, leading to the conclusion that there is an association in best response for patients in different arms of the trial.



#### 3.5. What factors are associated with tumor response (binary)?

##### 3.5.1. Logistic Model 1: Considering only Patient Demographics 

```{r, echo=FALSE}
# clean the data
dat$EOT_SUM_DIAMETER=as.numeric(dat$EOT_SUM_DIAMETER)

# drop NA
dat=dat %>% filter(BL_SUM_DIAMETER!="NA") %>% filter(EOT_SUM_DIAMETER!="NA")
dat=dat %>% filter(BESTRESP!="NA")  %>%
  mutate(
    age_group= ifelse(AGE > 60, "older than 60", "younger or equal to 60") )

dat_young=dat%>%filter(age_group=="younger or equal to 60")
dat_old=dat%>%filter(age_group=="older than 60")

# change y to numeric
dat$BESTRESP = as.numeric(dat$BESTRESP)

# in arm A, treatments echo DOX+EVO
# in arm B, treatments echo DOX only
mylogit <- glm(dat$BESTRESP ~ dat$ARMCD + age_group +RACE + ETHNIC, data = dat, family = "binomial")
summary(mylogit)

mylogit$coefficients %>% exp()
```

From the logistic regression model, we see that the p-values for ARMCDB, and RACEBLACK OR AFRICAN AMERICAN are significant, indicating that they are associated with tumor response. We can interpret them as:

1. If a patient belongs to Arm B, it will decrease the log odds of getting a tumor response by 0.554; so arm A (DOX+EVO) gives more tumor response Arm A is taken as reference.
2. If a patient's race is BLACK OR AFRICAN AMERICAN, it will decrease the log odds of getting a tumor response by 3.17. American Indian or Alaska Native is taken as reference.

We can also interpret the results using Odds ratio:

1. If a patient belongs to Arm B, the odds of getting a tumor response is 0.574; so arm A (DOX+EVO) gives more tumor response. Arm A is taken as reference.
2. If a patient's race is BLACK OR AFRICAN AMERICAN, the odds of getting a tumor response is 0.04. It is significantly less than the other groups. American Indian or Alaska Native is taken as reference.


##### 3.5.2. Logistic Model 2: Considering Patient Demographics and Treatment Baseline and Pathway

```{r, echo=FALSE}
# clean the data
dat$EOT_SUM_DIAMETER=as.numeric(dat$EOT_SUM_DIAMETER)

# drop NA
dat=dat %>% filter(BL_SUM_DIAMETER!="NA") %>% filter(EOT_SUM_DIAMETER!="NA")
dat=dat %>% filter(BESTRESP!="NA")  %>%
  mutate(
    age_group= ifelse(AGE > 60, "older than 60", "younger or equal to 60") )

dat_young=dat%>%filter(age_group=="younger or equal to 60")
dat_old=dat%>%filter(age_group=="older than 60")

# change y to numeric
dat$BESTRESP = as.numeric(dat$BESTRESP)
dat$HISTGRP = as.factor(dat$HISTGRP)
dat$STAGE = as.factor(dat$STAGE)
dat$PRADFL = as.factor(dat$PRADFL)
dat$PCHEMFL = as.factor(dat$PCHEMFL)

# in arm A, treatments echo DOX+EVO
# in arm B, treatments echo DOX only
mylogit <- glm(dat$BESTRESP ~ dat$ARMCD + age_group +RACE + ETHNIC + HISTGRP + STAGE + PRADFL + PCHEMFL, data = dat, family = "binomial")
summary(mylogit)
```

From the logistic regression model, we see that the p-values for ARMCDB, and RACEBLACK OR AFRICAN AMERICAN are significant, indicating that they are associated with tumor response. We can interpret them as:

1. If a patient belongs to Arm B, it will decrease the log odds of getting a tumor response by 0.626; so arm A (DOX+EVO) gives more tumor response. Arm A is taken as reference.
2. If a patient's race is BLACK OR AFRICAN AMERICAN, it will decrease the log odds of getting a tumor response by 3.47. American Indian or Alaska Native is taken as reference.

We can also interpret the results suing Odds ratio:

1. If a patient belongs to Arm B, the odds of getting a tumor response by 0.534; so arm A (DOX+EVO) gives more tumor response.
2. If a patient's race is BLACK OR AFRICAN AMERICAN, it will decrease the log odds of getting a tumor response by 3.47

#### 3.6. What factors are associated With percent change in sum of tumor diameters (continuous)? 

```{r, echo=FALSE}
model2 <- lm(dat$tumor_size_change ~ dat$ARMCD + AGE+RACE + ETHNIC, data = dat)
summary(model2)
```

No statistical significant factor. Also the model is not a good fitting model as most of the variables in our dataset are categorical.


### 4. Results 

#### 4.1. What conclusions can you make about the treatments’ effects on tumor response and percent change in sum of tumor diameters? 

Treatment effect on tumor response: According to the logistic regression function we did in Problem 3.1, if a patient belongs to arm B, it will decrease the log odds of getting a tumor response by 0.554; so arm A (DOX+EVO) gives higher tumor response.

Treatment effect on percent change in tumor diameters: According to the linear regression function we did in Problem 3.2, there is no significant factor associated with percent change in tumor diameters.


#### 4.2. Potential differential treatment effects in subgroups of patients (e.g., gender, race, age groups, etc.)

##### 4.2.1. For subgroup Gender, if treatment allocation (ARMCD) cause a change on the outcome (Best response)

```{r, echo=FALSE}
# split data by sex
dat_female=dat%>%filter(SEX=="F")
dat_male=dat%>%filter(SEX=="M")

# check treatment effects by Best response
female_logit <- glm(dat_female$BESTRESP ~ dat_female$ARMCD, data = dat_female, family = "binomial")
summary(female_logit)
# for female, the p-value of the treatment allocation (ARMCD) is not significant, so ARMCD does not cause a change on the tumor response in female. 

male_logit <- glm(dat_male$BESTRESP ~ dat_male$ARMCD, data = dat_male, family = "binomial")
summary(male_logit)
# for male, the p-value of the treatment allocation (ARMCD) is significant, so ARMCD does cause a change on the tumor response in male. If a male patient belongs to arm B, it will decrease the log odds of getting a tumor response by 0.61.
```

Logistic Regression Model of Response as a function of arm of treatment. The datasets are split into two: 1. Only Males 2. Only Females. We want to check if there is any significant change of outcomes when the only difference is the sex of the participants.

We see that males show statistical significance with treatment arms. The log odds of getting a positive outcome is reduced by .608 when males use Arm B. No such significance found for females. 


##### 4.2.2. For subgroup race, if treatment allocation (ARMCD) cause a change on the outcome (Best response)

```{r, echo=FALSE}
# split data by the top 3 representative race
dat_white=dat%>%filter(RACE=="WHITE")
dat_black=dat%>%filter(RACE=="BLACK OR AFRICAN AMERICAN")
dat_asia=dat%>%filter(RACE=="ASIAN")


# check treatment effects in Best response
white_logit <- glm(BESTRESP ~ ARMCD, data = dat_white, family = "binomial")
summary(white_logit)
# for white race, the p-value of the treatment allocation (ARMCD) is significant, so ARMCD does cause a change on the tumor response in female. 
# If a white patient belongs to arm B, it will decrease the log odds of getting a tumor response by 0.51

black_logit <- glm(BESTRESP ~ ARMCD, data = dat_black, family = "binomial")
summary(black_logit)
# for black race, the p-value of the treatment allocation (ARMCD) is not significant, so ARMCD does not cause a change on the tumor response in black race.

asia_logit <- glm(BESTRESP ~ ARMCD, data = dat_asia, family = "binomial")
summary(asia_logit)
# for Asian race, the p-value of the treatment allocation (ARMCD) is not significant, so ARMCD does not cause a change on the tumor response in asia race.

```

Logistic Regression Model of Response as a function of arm of treatment. The datasets are split into three: 1. White Participants 2. Black Participants 3. Asian Participants. We want to check if there is any significant change of outcomes when the only difference is the race of the participants.

We see that white and black participants show statistical significance with treatment arms. For white participants, the log odds of getting a positive outcome is reduced by .519 when white participants use Arm B. For black participants, the log odds of getting a positive outcome is reduced by .319 when black participants use Arm B.


##### 4.2.3. For subgroup Age groups, if treatment allocation (ARMCD) cause a change on the outcome (Best response)

```{r, echo=FALSE}
# split the data by 2 age groups
dat <- dat %>%
  mutate(
    age_group= ifelse(AGE > 60, "older than 60", "younger or equal to 60") )

dat_young=dat%>%filter(age_group=="younger or equal to 60")
dat_old=dat%>%filter(age_group=="older than 60")

# check treatment effects in Best response
young_logit <- glm(BESTRESP ~ ARMCD, data = dat_young, family = "binomial")
summary(young_logit)
# for people younger or equal to 60, the p-value of the treatment allocation (ARMCD) is significant, so ARMCD does cause a change on the tumor response. If a patient who is younger or equal to 60 and belongs to arm B, it will decrease the log odds of getting a tumor response by 0.78.

old_logit <- glm(BESTRESP ~ ARMCD, data = dat_old, family = "binomial")
summary(old_logit)
# for people older than 60 male, the p-value of the treatment allocation (ARMCD) is not significant, so ARMCD does not cause a change on the tumor response in older people.
```

Logistic Regression Model of Response as a function of arm of treatment. The datasets are split into two: 1. Participants less than or equal to 60 2. Participants greater than 60. We want to check if there is any significant change of outcomes when the only difference is the age group of the participants.

We see that participants younger or equal to 60 show statistical significance with treatment arms. The log odds of getting a positive outcome is reduced by .78 when participants use Arm B. No statistical significance for participants older than 60.


### 5. Conclusions

According to the analysis, we can drive to the conclusion that drug combination group (DOX + EVO) gives more tumor  response than mono (DOX) treatment. However, there is no significant factor associated with percent change in tumor diameters by using linear regression. To specify, if a patient is male, or belongs to white race, or younger or equal to 60 years old, the individual would experience more tumor response if receiving treatment DOX + EVO, while keeping other factors fixed.


### 6. Limitations

There still would be some limitations for this study. Firstly, a significantly higher proportion of white ethnicity participants is observed. This can lead to biased outcomes. Also, the patients' stages at diagnosis are not equally allocated in the two arms. There are more patients in stage III at diagnosis. This may be another factor leading to biased outcomes.



### 7. References

1. Verschoor, Arie Jan, et al. "Survival of soft tissue sarcoma patients after completing six cycles of first-line anthracycline containing treatment: an EORTC-STBSG database study." Clinical sarcoma research 10.1 (2020): 1-9.
2. Tseng, William W., Behrang Amini, and John E. Madewell. "Follow‐up of the soft tissue sarcoma patient." Journal of surgical oncology 111.5 (2015): 641-645.
3. Soft tissue sarcoma-patient version. National Cancer Institute. (n.d.). Retrieved December 8, 2022
